# Axioms of Kolmogorov

(1) **Non-negativity Axiom**

$$
\text{For any event A in our sample space, } \mathbb{P}(A)≥0
$$

(2) **Unity Axiom**

$$
\text{Normalization, also called the assumption of unit measure,}\\ \text{defines the probability of the sample space }
\Omega \text{ as } \mathbb{P}(\Omega)=1
$$

(3) $\sigma$**-additivity Axiom**

$$
\text{Any countable sequence of disjoint (mutually exclusive) sets satisfies} 
\\ 
\mathbb{P}(\bigcup_{i=1}^{\infty}E_{i}) = \sum_{n=1}^{\infty}\mathbb{P}(E_{i})
\qquad\text{where }\ E_i\cap E_j = \varnothing, \quad \forall i\not=j
$$

**Additional Notes**

-   These axioms allow us to derive the theorems and properties that we frequently come across in the study of probability.

-   Some authors tend to just use finite additivity as the third axiom.

-   Stated in two axioms:

$(1) \qquad \text{Axiom 1 }\mathbb{P(\varnothing) =0, P(\Omega)=1}, \\ (2) \qquad\sigma-\text{additivity}$

## Implied consequences of the Axioms

Demonstrating these immediate consequences illustrates the power of the third axiom, and its interaction with the remaining two axioms.

------------------------------------------------------------------------

$$
\text{ Probability of the empty set} 
\\ 
\mathbb{P}(\varnothing) = 0
\tag{Theorem 2.1}
$$

------------------------------------------------------------------------

$$
\text{Finite Additivity}
\\
\mathbb{P}(\bigcup_{n=1}^{N} E_i)=\sum_{n=1}^{N}\mathbb{P}(E_i)
\tag{Theorem 2.2}
$$

------------------------------------------------------------------------

$$
\text{Monotonicity}
\\
A\subseteq B \text{ then }\mathbb{P}(A)\le\mathbb{P}(B)
\tag{Theorem 2.3} 
$$

-   Understand that $A\subseteq B$ means that if a occurs, then B must occur.

------------------------------------------------------------------------

$$
\text{The numeric bound}
\\
0\le\mathbb{P}(E)\le1, \qquad\forall E\in F
\tag{Theorem 2.3} 
$$

------------------------------------------------------------------------

$$
\text{The complement rule} 
\\ \mathbb{P}(E^c) = \mathbb{P}(\Omega-E) = 1 - \mathbb{P}(E)\tag{Theorem 2.4}
$$

------------------------------------------------------------------------

$$
\text{Inclusion-Exclusion Principal} 
\\
P(A\cup B)=P(A) + P(B)-P(AB)
\tag{Theorem 2.5}
\\
$$

------------------------------------------------------------------------

$$
\text{Generalized Inclusion-Exclusion Principle} 
\\
\mathbb{P}(\bigcup_{i=1}^n A_i)  = \sum_{i=1}^{n} (-1)^{i+1} 
\\
\sum_{\Omega \subseteq} \mathbb{P}(\bigcup_{i \in \Omega}  A_{i})
\tag{Theorem 2.5}
$$

------------------------------------------------------------------------

$$
\text{Equally Likely Theorem} \\ \mathbb{P}(A)=\frac{N(A)}{N}\tag{Theorem 2.7}
$$

-   This is the naive definition of probability

------------------------------------------------------------------------

$$
\text{No Name Theorem}
\\ 
\mathbb{P}(A) = P(AB) + P(AB^c)
\tag{Theorem 2.8}
$$

------------------------------------------------------------------------

$$
\text{Boole's Inequality}
\\ 
\mathbb{P}(\bigcup_{i=1}^{\infty}E_{i}) \le \sum_{n=1}^{\infty}\mathbb{P}(E_{i})
$$

## About Sets, Sample Spaces, and events

-   Sets are collections of distinct elements.
-   A sample space is the set of disjoint, **collectively exhaustive** outcomes taken at a determined level of granularity.
-   A sample space can be described as finite or infinite, discrete or continuous.

**Sample Space Notation**: A sample space can be described with notation below, the long bar, \| is read as "such that"

$$
A\cup B=\{x| x∈A \cup x∈B\}
\\
A^c=\{x|x\not\in A\}
$$

### Set Notation and LaTex code.

```{r, echo=FALSE}
library(knitr)
library(kableExtra)

table_data <- data.frame(
  Symbol = c("$\\cup$", "$\\cap$", "$\\in$", "$\\varnothing$", 
             "$\\Omega$, $S$", "$\\omega$, $s$", "$\\subset$", "$\\subseteq$", 
             "$\\bigcup$", "$\\bigcap$"),
  
  Name = c("Union", 
           "Intersection",
           "Belongs to", 
           "Null set", 
           "Universal set", 
           "Atom, singleton", 
           "Proper subset",
           "Subset, S is contained within A",
           "infinitary union, generalized union, unified union",
           "infinitary intersection, generalized intersection, unified intersection cap"),
  
  LaTeX = c("\\\\cup", "\\\\cap", "\\\\in", "\\\\varnothing",
            "\\\\Omega", "\\\\omega", "\\\\subset", "\\\\subseteq", "\\\\bigcup", "\\\\bigcap")
  
  )

kable_styling(kable(table_data, format = "html", escape = FALSE))
```

### Some Set Definitions

```{r, echo=FALSE}
library(knitr)
library(kableExtra)

table_data <- data.frame(
  Name = c("Certainty", "Impossibility", "Mutual Exclusivity", "Complement", "Sum Rule", "Difference Rule"),
  
  Definition = c("The probability of the sample space (the set of all possible outcomes) is
                 1, representing absolute certainty",
                 "The probability of an impossible event is 0",
                 "Two events A and B cannot occur simultaneously, in terms of probability",
                 "The complement of event A is self-defined",
                 "The probability of the union of two mutually exclusive events A and B",
                 "The probability of the difference between two events A and B"
  ),
  Rule = c("$P(\\Omega) = 1$",
           "$P(\\varnothing) = 0$",
           "$P(A \\cap B) = 0$",
           "$P(A^c) = 1 - P(A)$",
           "$P(A \\cup B) = P(A) + P(B)$",
           "$P(A - B) = P(A) - P(A \\cap B)$"
  )
)

kable_styling(kable(table_data, format = "html", escape = FALSE))
```

> Note that Impossibility and Certainty are both themselves axioms

### Relational Laws

```{r, echo=FALSE}
library(knitr)
library(kableExtra)

table_data <- data.frame(
  Name = c("Commutative", "Associative", "Distributive", "Demorgans First Law", "Demorgans Second Law"),
  
  Law = c("$AB = BA$",
          '$(AB)C = A(BC)$',
          "$AB \\cup C = (A \\cup C)(B \\cup C)  ,\\quad A(B\\cup C) = AB \\cup AC$",
          "$(A \\cup B)^c = A^cB^c$",
          "$(AB)^c = A^c \\cup B^c$")
)

kable_styling(kable(table_data, format = "html", escape = FALSE))
```

    [Future sections coming soon]

    Proofs for consequences

    About Probability Measures

    Continuity of Probability Functions
