# Combinatorics

## Counting Principal

The Fundamental Counting Principle, I also referred to as the generalized counting principal, the rule of product, the multiplication principal, and the multiplication rule. The rule states that the number different possible outcomes for a sequence of $n$ stages is the product of the number of different ways each event can occur.

$$
\text{Let }E_1, E_2, ..., E_k \text{ be sets of events, where each set has }n_1,n_2, ...,n_k \text{ respective elements.}
\\
\text{Then there are }n_1×n_2×n_3×···×n_k \text{ combinations in which we, first,}
\\
\text{choose an element of }E_1,
\text{then choose element of }E_2, \text{ ... , ending with } E_k.
$$ - E_k-1 ,E_k

[Combinatorial principles - Wikipedia](https://en.wikipedia.org/wiki/Combinatorial_principles)

The counting principal can be visualized with a tree graph.

There are some immediate rules that can be followed from the counting principal.

```{r, echo=FALSE}
library(knitr)
library(kableExtra)


table_data <- data.frame(Rule = c("Permutations, w/ Replacement",
                                  "Permutations w/o Replacement",
                                  "Combinations, w/o Replacement",
                                  "Combinations w/ Replacement",
                                  "Number of Possible Subsets in a Set"),
                         
                         Equation= c("$n^k$",
                                     "$\\frac{n!}{(n-r)!}$",
                                     "$\\frac{n!}{r!(n-r)!}$",
                                     "$\\binom{n+k-1}{k}$",
                                     "$2^n$"))

kable_styling(kable(table_data, format = "html", escape = FALSE))
```

## Permutations

A permutation is the number of ways we can order $n$ distinct objects. From the generalized counting principle we have $n$ ways to make the first selection. Then we will have one fewer way to make our next selection, and so on... until we only have one option for the last selection.

This results in the equation

$$
(n)(n−1)(n−2)···(1)=n!
$$

We use the notation $_nP_r$ to denote the number of permutations of a set $A$ containing $n$ elements taken $r$ at a time where $(1 ≤ r ≤ n)$

The number of r-element permutations of a set containing n objects is given by

$$
_nP_r = \frac{n!}{(n-r)!}
$$

(note: to avoid division by zero when $n=r$ we define $0! = 1$)

To derive the formula for $_nP_r$ we consider the number of choices for each selection.

Since $A$ has $n$ elements, the number of choices for the first object in the $r$-element permutation is $n$. For the next selection there will be $n-1$ options, followed by $n-2$ options for the selection that follows after that, we do this until our last selection, **which is the** $r$**th event (or object)** which has $n − (r − 1)$ possible choices.

```{r, echo=FALSE}
cat("_____    _____    _____  ...   _____  \n",
    "(n)     (n-1)    (n-2)  ...  (n-(r-1))  ")
```

**We will always be accounting for the ways you can permeate** $r$ object, the result of $n - (r - 1)$ is the number of objects taken into account.

This leads us to the equation

$$
_nP_r =n(n−1)(n−2)···(n−r+1)
$$

Aside: Permutations with $_nP_n$ equate to

$$
_nP_n =n(n−1)(n−2)···(n−n+1)=n!
$$

Manipulating the equation $n(n−1)(n−2)···(n−r+1)$ leads us to the general formula

$$
_nP_r = \frac{n!}{(n-r)!}
$$

### Permutations with Replacement

Using the counting principal we see that the number of permutations when we use replacement sampling is equal to

$$
n^k = (n)(n)....(n_k)
$$

### Permutations with distinguished elements

The formula above is valid only if all of the objects of the set are distinguishable from each other. If there are repeated elements then we must divide to cancel out the permutations that are a result of repeated elements

$$
\frac{n!}{n_1! × n_2 ! × · · · × n_k!}
$$

Consider the names **Brungard** and **Karapanian**. If for some reason we wanted to know the possible ways to rearrange these respective letters:

1.  Brungard

$$
\frac{8!}{2!} = 40320
$$

2.  Karapanian

Here are ten letters, but only 6 distinct letters, 4 are the letter a and 2 are the letter n

$$
\frac{10!}{4! *2!}= 75600
$$

While there are still more ways to permiate the longer last name, consider if you did not account for repeated characters... the overcount is over three million..

```{r}
factorial(10)/(factorial(4)*factorial(2))

factorial(10) 
```

## Combinations

we have for $r \le n$

$$
\binom{n}{r}=\frac{n!}{r!(n-r)!}
$$

We are calculating the number of $r$-element subsets from a $n$-element set, where order matters. This is the number of all r-element combinations of n objects, where we only use valid non-negative integers (e.g. 0,1,2, ...) where $r \le n$.

[add derivation, xr! = npr]

### Combinations with replacement

$$
\binom{n+k-1}{k}
$$

------------------------------------------------------------------------

### Useful Relations

**Choosing all and choosing nothing**

There is only one combination to choose all elements in a set. Similarly, there is only one way to choose nothing.

$$
\binom{n}{n}=1
\qquad \& \qquad  
\binom{n}{0}=1
$$ Both relations can be easily be shown algebraically

$$
\binom{n}{n}=\frac{n!}{n!(n-n)!}=1
$$

------------------------------------------------------------------------

**Choosing 1 from n objects is the same as choosing n-1 fron n objects**

$$
\binom{n}{1}=\binom{n}{n-1}
$$

consider this generalization

where $0 \le r \le n$

$$
\binom{n}{r}=\binom{n}{n-r}
$$

and

$$
\binom{n+1}{r}=\binom{n}{r}+\binom{n}{r-1}
$$

## Number of Sets in a Subset

The set of all subsets of A is called the **power set** of A.

We can calculate the number of subsets in a set as $2^n$, because each element can be either included or excluded (hence the two options)

    [write more thoughts about this and how to verify algebraically or using combinatorics]

    ## Examples

    [coming soon]

    ## Practice Problems

    [coming soon]

    example 2.22 and binomial expansion
